---
title: "Reproducible_Proj1"
author: "Tyler Shirley"
date: "8/4/2020"
output:
  html_document: 
    keep_md: yes
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read Dataset

Read the original activity dataset and load all necessary packages to process the data.

Original data is scrubbed of NA values. Then put date and time values into the dataframe.
```{r}
activity_unclean <- read.csv("activity.csv")

#Load in packages
library(ggplot2)
library(lubridate)
library(plyr)
library(dplyr)

#remove the NA values & make date data type
activity_unclean$date <- as.Date(as.character(activity_unclean$date))
activity <- na.omit(activity_unclean)
activity$day <- weekdays(activity$date)

step_day <- aggregate(steps ~ date, activity, FUN = sum)
```

## Histogram of Daily Steps

Create a histogram by looking at the frequency of steps by date.
```{r}
hist(step_day$steps, 
     xlab = "average steps",
     main = "Histogram of Steps",
     col = "red")
```

## Mean and Median of steps
```{r}
mean_median_steps <- summary(step_day$steps)
mean_median_steps
```

## Time Series Number of Steps

Group the steps by the intervals taken. Then use the ggplot functions to construct a graph of Average steps by Interval.

```{r}
step_int <- aggregate(steps ~ interval, activity, FUN = mean)
tim_ser <- ggplot(step_int, aes(x = interval, y = steps)) + 
  geom_line() + 
  labs(title = "Average Steps Per Interval", x = "Interval", y = "Average Steps") +
  theme(plot.title = element_text(hjust = 0.5))

tim_ser
```

## Max Step intervals

find the interval with the maximum number of steps

```{r}
max_step <- max(step_int$steps)
biggest_interval <- step_int[which.max(step_int$steps), 1]

Step_max_tbl <- data.frame("Max Steps" = max_step,
                           "Biggest interval" = biggest_interval)

Step_max_tbl
```

##Number of NA in Active Dataset

```{r}
sum_active_na <- sum(is.na(activity_unclean))
sum_active_na
```

## Replace NA with Averages from Intervals

Create a logical vector of the NA values from active_na. Then create a vector of averages based on the intervals available.
```{r}
df_active_na <- activity_unclean
active_na <- is.na(df_active_na$steps)
averages <- tapply(df_active_na$steps, df_active_na$interval, mean, na.rm = TRUE)

df_active_na <- df_active_na %>% 
  mutate(steps = replace(steps, active_na, averages))
  
new_df_active <- tapply(df_active_na$steps, df_active_na$date,sum, na.rm = TRUE)
hist(new_df_active,
     main = "Histogram of New Averages",
     xlab = "Steps By Date",
     ylab = "Instances of Steps Per Day",
     col = "blue")

Mean_new = mean(new_df_active)
Median_new = median(new_df_active)

Replaced_Data <- data.frame("New Mean" = Mean_new,
                            "New Median" = Median_new)
Replaced_Data

```

## Weekends vs Weekday

Find the breakdown of weekend steps vs weekday steps. First make new columns to show what is a weekday vs wat is a weekend. Then create new dataframes with only Weekend or Weekday data. That is then aggregated with the averages to show the means across the Weekday an Weekend Dataframe. 
```{r}
df_active_na$day <- weekdays(df_active_na$date)

df_active_na$dow <- ifelse(df_active_na$day %in% c("Saturday", "Sunday"),
                            "Weekend",
                            "Weekday")

Weekend_int <- subset(df_active_na, df_active_na$dow == "Weekend")
Weekend_int <- aggregate(steps ~ interval, Weekend_int, FUN = mean)

Weekday_int<- subset(df_active_na, df_active_na$dow == "Weekday")
Weekday_int <- aggregate(steps ~ interval, Weekday_int, FUN = mean)

par(mfrow = c(2,1))
  plot(Weekday_int$interval, Weekday_int$steps, "l",
       main = "Weekday Steps",
       xlab = "Weekday Interval",
       ylab = "Weekday Steps")
  plot(Weekend_int$interval, Weekend_int$steps, "l",
       main = "Weekend Steps",
       xlab = "Weekend Interval",
       ylab = "Weekend Steps")

```